# Karen's Week 5 Work

## Instructions

The following exercises are a test in disguise. What do you think about the coding? Can you think of any improvements to the following?

## Preparations

### Installing packages

```{r}
library(tidyverse)
library(ggplot2)
```

### Loading datasets

```{r}
aq = readRDS("anscombe_quartet.rds")
dd = readRDS("datasaurus_dozen.rds")
```

## PROBLEM 1

Visualizing your data is important! Summary statistics can be highly misleading, and simply plotting the data can reveal a lot more! Lets look at the Anscombe’s Quartet data. There are four different data sets.

*Anscombe, F. J. (1973). "Graphs in Statistical Analysis". American Statistician. 27 (1): 17–21. <doi:10.1080/00031305.1973.10478966>. JSTOR 2682899.*

### Data exploration

```{r}
str(aq)
```

*Q: What does the function str() do?*

A: It displays the internal structure of the dataset.

```{r}
aq %>% group_by(dataset) %>% 
  summarise( mean_x =mean(x), 
             mean_y = mean(y), 
             min_x = min(x), 
             min_y = min(y), 
             max_x = max(x), 
             max_y = max(y), 
             crrltn = cor(x, y) )
```

*Q: what do the summary statistics tell us about the different datasets?*

A: All datasets have identical mean values of x and y, and highly similar correlations, but have distinct min and max x and y values.

### Data visualisation

```{r}
plot1 <- ggplot(aq, aes(x=x,y=y)) + 
  geom_point() +
  geom_smooth(method = "lm",formula = "y ~ x") + facet_wrap(~dataset)

ggsave("plot1.png", width = 20, height = 20, units = "cm")
```

*Q: what do the plots tell us about the different datasets?*

A: it visualises the relationship between x and y values in each dataset

*Q: describe the relationship between x and y in the different datasets.*

A: 1. fairly linear 2. quadratic 3. very linear with what looks like an outlier 4. all data points except 1 have the same x value and different y values

*Q: would linear regression be an appropriate statistical model to analyse the x-y relationship in each dataset?*

A: It is suitable for 3 and potentially for 1, but not for 2 and 4

*Q: what conclusions can you draw for the plots and summary statistics?*

A: summary statistics don't accurately capture characteristics of the dataset, so visualisation is very important.

## PROBLEM 2

### Data exploration

```{r}
summary(dd)
```

*Q: how many rows and columns does the datasaurus_dozen file have?*

A: 2 columns and 1846 rows

### Data visualisation

```{r}
ggplot(dd, aes(x=x,y=y)) + geom_point() + facet_wrap(~dataset)
```

*Q: calculate the correlations and summary statistics for x and y in all datasets:*

```{r}
# A:
dd %>% 
  group_by(dataset) %>% 
  summarise( mean_x = mean(x), 
             mean_y = mean(y), 
             min_x = min(x), 
             min_y = min(y), 
             max_x = max(x), 
             max_y = max(y),
             crrltn = cor(x, y) )
```

*Q: Plot the relationships between x and y in each dataset including the line of best fit.*

```{r}
# A:
plot2 <- ggplot(dd, aes(x=x,y=y)) + geom_point() + geom_smooth(method = "lm",formula = "y ~ x") + facet_wrap(~dataset)
ggsave("plot2.png")
```

*Q: what conclusions can you draw for the plots and summary statistics?*

A: plots represent different images, x and y vales are not correlated in a linear manner in any of the datasets, which is not represented in the summary statistics.
